A holographic approach to the six-dimensional superconformal index

### 2022-10-14 fivebranes/intro.tex
merge

A principal goal of the present paper is to initiate the study of the factorization algebras of observables associated to the minimal twists of the six-dimensional~$\cN=(2,0)$ superconformal field theories. The full factorization structure is extremely rich, and in this paper, we only discuss a small part of it: the costalk at a point as a vector space---otherwise known as the space of local operators at the point. 
However, this information is enough to extract familiar quantities like the superconformal index.

@@

To describe the kinds of theories on either side of the correspondence, we should comment on what it means to twist a theory of supergravity. We will provide more details in \S \ref{s:twisted}. Since the action of supersymmetry is gagued in theories of supergavity, the aforementioned twisting procedure does not quite make sense. Instead, Costello and Li define twisted supergravity to be the theory in perturbation theory around a background where the bosonic ghost for local supertranslations takes a nonzero nilpotent VEV. Coupling the worldvolume theory of a brane to such a background has the effect of twisting the worldvolume theory in the usual sense.

@@

\subsection{The fivebrane decomposition}
We will argue that the value of the factorization algebra associated to the worldvolume theory on a finite number of minimally twisted fivebranes on an arbitrary open set is a particular piece of the $!$-dual of the factorization algebra associated to the bulk gravitational theory.
After taking a certain classical limit, this argument relies on a presentation of the $!$-dual as
\beqn
(\Obs_{grav}|_{Y})^{!} \simeq \cF_{-1} \otimes \cF_0 \otimes \cF_1 \otimes \cdots 
\eeqn
for some factorization algebras $\cF_{-1},\cF_0, \cF_{1},\ldots$ which are defined using a certain weight decomposition of the bulk gravitational theory.
In fact, this decomposition is intimately related to a certain decomposition of the exceptional simple super Lie algebra $E(5|10)$ that we have already mentioned plays an important role in the minimal twist of eleven-dimensional supergravity. This decomposition is a natural lift of the grading on $\operatorname{Ham}(\C^{2})$ induced from taking the associated graded with respect to the order filtration on $\operatorname{Diff} (\C^{2})$. The strange indexing conventions will be explained in \S\ref{s:fact}.

@@

A related theory is the minimal twist of the six-dimensional $\cN=(2,0)$ superconformal theory associated to a Lie algebra of type $A_{N-1}$.
We denote the corresponding factorization algebra by $\Obs_{A_{N-1}}$ for now. 
This is obtained from the worldvolume theory simply by throwing away the modes propagating transverse to the brane, which in our presentation above corresponds to stripping off the first factor $\cF_{-1}$. 
Thus, at the level of factorization algebras we have a similar proposed decomposition
\beqn
\label{eqn:AN-1}
\Obs_{A_{N-1}} = \cF_0 \otimes \cF_1 \otimes \cdots \otimes \cF_{N-2} .
\eeqn

We want to emphasize that these expressions only hold after taking this classical limit and taking some truncation of the differential present on the left-hand side.\footnote{An instructive avatar of this decomposition to keep in mind is the description of the $W_N$ vertex algebra as being generated by a collection of operators of spins $2,3,\ldots, N$.
Only in the classical limit does the $W_N$ vertex algebra decompose as a pure tensor product.}
This decomposition as a tensor product of factorization algebras will break down at the quantum level.
Nevertheless, the description provides a very effective way to compute certain protected quantities like the superconformal index.

@@

If we are working simply on the worldvolume $\C^3 \cong \R^6$, our description in \eqref{eqn:finiteTensor} posits that after taking the classical limit the decomposition of factorization algebras associted to a stack of $N$ twisted fivebranes induces a decomposition at the level of local operators
\beqn
\cF_{-1}(0) \otimes \cF_0(0) \otimes \cdots \otimes \cF_{N-2}(0) ,
\eeqn
where $\cF_k(0)$ stands for the algebra of local operators associated to the factorization algebra $\cF_k$.
Forgetting about the differentials for the time being, the local operators $\cF_k(0)$ are given as the free algebra on some vector space $V_k$ of linear local opeartors.
Thus, to compute the index of local operators of $\Obs_{brane}(N)$ it suffices to compute the index of each of the $V_k$---this is the single particle index---and then apply the plethystic exponential.
For now, denote by $g_k({\bf x})$ the index of the vector space $V_k$.
This quantity is directly computable from the gravitational side.

@@

Even stronger evidence would involve the factorization algebras $\Obs_{N-branes}$. We will outline some expectations in later sections but leave a full treatment to future work.

@@

We propose similar statements at the level of the minimal twist. 
We fix a complex surface $S$ together with a pair of complex vector bundles $L,W \to S$ of ranks $1$ and $2$ respectively, satisfying the condition that $X = \text{Tot} (L \oplus W)$ is a Calabi--Yau 5-fold.
The twisted eleven-dimensional supergravity lives on the eleven-manifold $\R \times X$.

@@

As in the abelian case, we may highlight some compatibilities with the usual AGT correspondence. 
For a particular choice of the input geometric data, the same Maurer-Cartan element in $\lie{osp}(6|2)$ deforms the associative algebra $\pi_{*}\Obs_{A_{1}}|_{\R_{>0}}$ to the Virasoro algebra $U(\operatorname{Vir})$. We again expect that for a judicious choice of $X$, the Hilbert space deforms to the equivariant cohomology of the moduli of rank 2 instantons on $\C^{2}$, and the action compatibly deforms to the action in the rank 2 case of the AGT correspondence.

A result in the main body of the paper may be viewed as a check of these conjectures in the case where $Z = \operatorname{Tot}(\cO(-1)\to \C\PP^{2})$ and $W = K^{1/2}_{Z}\otimes \C^{2}$. In this case, the unit circle bundle in $L$ may be identified with the Hopf fibration $S^{5}\to \C\PP^{2}$. In \cite{Kim:2013nva}, the superconformal index of the 6d $\cN=(2,0)$ theory is computed by identifying it with the $S^{1}\times S^{5}$ partition function, and reducing along the Hopf fiber of $S^{5}$ to further identify it with the superconformal index of 5d $\cN=2$ gauge theory on $\C\PP^{2}$. The latter quantity is computed in terms of the Nekrasov instanton partition function, and is almost manifestly the character of $\cO \left ( \operatorname{Higgs} _{SL(2)}(S, W)\right)$. The fact that we recover the same expressions as characters of the algebras of local operators $\Obs_{SL(2)}(0)$ corroborates the claim that $ \cO \left ( \operatorname{Higgs} _{SL(2)}(S, W)\right)$ is the vacuum module of $\pi_{*}\Obs_{SL(2)}|_{\R_{>0}}$.

>>>>

A principal goal of the present paper is to initiate the study of the factorization algebras of observables associated to the minimal twists of the six-dimensional~$\cN=(2,0)$ superconformal field theories. The full factorization structure is extremely rich, and in this paper we only discuss a small part of it, namely the costalk at a point as a vector space.
This is otherwise known as the space of local operators at the point. 
However, this information is enough to extract familiar quantities like the superconformal index.

@@

To describe the kinds of theories on either side of the correspondence, we should comment on what it means to twist a theory of supergravity. We will provide more details in \S \ref{s:twisted}. Since the action of supersymmetry is gagued in theories of supergavity, the aforementioned twisting procedure does not quite make sense. Instead, Costello and Li define twisted supergravity to be the theory in perturbation theory around a background where the bosonic ghost for local supertranslations takes a nonzero nilpotent vacuum expectation value. 
Coupling the worldvolume theory of a brane to such a background has the effect of twisting the worldvolume theory in the usual sense.

@@

\subsection{The fivebrane decomposition}
We will argue that the value of the factorization algebra associated to the worldvolume theory on a finite number of minimally twisted fivebranes on an arbitrary open set is a particular piece of the $!$-dual of the factorization algebra associated to the bulk gravitational theory.
After taking a certain limit, this argument relies on a presentation of the $!$-dual as
\beqn
(\Obs_{grav}|_{Y})^{!} \simeq \cF_{-1} \otimes \cF_0 \otimes \cF_1 \otimes \cdots 
\eeqn
for some factorization algebras $\cF_{-1},\cF_0, \cF_{1},\ldots$ which are defined using a certain weight decomposition of the bulk gravitational theory.
In fact, this decomposition is intimately related to a certain decomposition of the exceptional simple super Lie algebra $E(5|10)$ that we have already mentioned plays an important role in the minimal twist of eleven-dimensional supergravity. This decomposition is a natural lift of the grading on $\operatorname{Ham}(\C^{2})$ induced from taking the associated graded with respect to the order filtration on $\operatorname{Diff} (\C^{2})$. The strange indexing conventions will be explained in \S\ref{s:fact}.

@@

A related theory is the minimal twist of the six-dimensional $\cN=(2,0)$ superconformal theory associated to a Lie algebra of type $A_{N-1}$.
We denote the corresponding factorization algebra by $\Obs_{A_{N-1}}$ for now. 
This is obtained from the worldvolume theory simply by throwing away the modes propagating transverse to the brane, which in our presentation above corresponds to stripping off the first factor $\cF_{-1}$. 
Thus, at the level of factorization algebras we have a similar proposed decomposition
\beqn
\label{eqn:AN-1}
\Obs_{A_{N-1}} (U) \simeq \cF_0 (U) \otimes \cF_1 (U) \otimes \cdots \otimes \cF_{N-2} (U).
\eeqn

We want to emphasize that these expressions only hold after taking this classical limit and taking some truncation of the differential present on the left-hand side.\footnote{An instructive avatar of this decomposition to keep in mind is the description of the $W_N$ vertex algebra as being generated by a collection of operators of spins $2,3,\ldots, N$.
Only in the classical limit does the $W_N$ vertex algebra decompose as a pure tensor product.}
This decomposition as a tensor product of graded vector spaces will break down at the quantum level and when we take into account the factorization algebra structure.
Nevertheless, the description provides a very effective way to compute certain protected quantities like the superconformal index.

@@

If we are working simply on the worldvolume $\C^3 \cong \R^6$, our description in \eqref{eqn:finiteTensor} posits that after taking the classical limit the decomposition of factorization algebras associted to a stack of $N$ twisted fivebranes induces a decomposition at the level of local operators
\beqn
\cF_{-1}(0) \otimes \cF_0(0) \otimes \cdots \otimes \cF_{N-2}(0) ,
\eeqn
where $\cF_k(0)$ stands for the space of local operators associated to the factorization algebra $\cF_k$.
Forgetting about the differentials for the time being, the local operators $\cF_k(0)$ are given as the free symmetric algebra on some vector space $V_k$ of linear local opeartors.
Thus, to compute the supersymmetric index of the space of local operators of $\Obs_{N-branes}$ it suffices to compute the index of each of the $V_k$---this is the single particle index---and then apply the plethystic exponential.
For now, denote by $g_k({\bf x})$ the index of the vector space $V_k$.
This quantity is directly computable from the gravitational side.

@@

Even stronger evidence would involve understanding what the gravitational side can say about the factorization algebras $\Obs_{N-branes}$. 
We will return to a treatment of this in future work.

@@

We propose similar statements at the level of the minimal twist. 
We fix a complex surface $S$ together with a pair of complex vector bundles $L,W \to S$ of ranks one and two respectively, satisfying the condition that $X = \text{Tot} (L \oplus W)$ is a Calabi--Yau fivefold.
The twisted eleven-dimensional supergravity lives on the eleven-manifold $\R \times X$.

@@

As in the abelian case, we may highlight some compatibilities with the usual AGT correspondence. 
For a particular choice of the input geometric data, the same Maurer-Cartan element in $\lie{osp}(6|2)$ will deform the associative algebra $\pi_{*}\Obs_{A_{1}}|_{\R_{>0}}$ to the Virasoro algebra $U(\operatorname{Vir})$. We again expect that for a judicious choice of $X$, the Hilbert space deforms to the equivariant cohomology of the moduli of rank 2 instantons on $\C^{2}$, and the action compatibly deforms to the action in the rank 2 case of the AGT correspondence.

A result in the main body of the paper may be viewed as a check of these conjectures in the case where $Z = \operatorname{Tot}(\cO(-1)\to \C\PP^{2})$ and $W = K^{1/2}_{Z}\otimes \C^{2}$. In this case, the unit circle bundle in $L$ may be identified with the Hopf fibration $S^{5}\to \C\PP^{2}$. In \cite{Kim:2013nva}, the superconformal index of the six-dimensional $\cN=(2,0)$ theory is computed by identifying it with the $S^{1}\times S^{5}$ partition function, and reducing along the Hopf fiber of $S^{5}$ to further identify it with the superconformal index of five-dimensional $\cN=2$ gauge theory on $\C\PP^{2}$. The latter quantity is computed in terms of the Nekrasov instanton partition function, and is almost manifestly the character of $\cO \left ( \operatorname{Higgs} _{SL(2)}(S, W)\right)$. 
The fact that we recover the same expression as the character of the space of local operators $\Obs_{A_1}(0)$ corroborates the claim that $ \cO \left ( \operatorname{Higgs} _{SL(2)}(S, W)\right)$ is the vacuum module of $\pi_{*}\Obs_{A_1}|_{\R_{>0}}$.
